---
title: "p8105_hw5_az2852"
output: github_document
---

```{r setup, message = FALSE}
library(tidyverse)
library(rvest)
library(broom)
library(stringr)
```

# Problem 1
Suppose you put n people in a room, and want to know the probability that at least two people share a birthday. For simplicity, we’ll assume there are no leap years (i.e. there are only 365 days) and that birthdays are uniformly distributed over the year (which is actually not the case).

```{r}
set.seed(123)

sample = sample(1:365, size = 10, replace = TRUE)

length(unique(sample)) < 10
```

write a function
```{r}
bday_sim = function(n) {
  
  sample = sample(1:365, size = n, replace = TRUE)

duplicates = length(unique(sample)) < n

return(duplicates)
}

bday_sim(40)
```
run this function 10000 times for each group size between 2 and 50. For each group size, compute the probability that at least two people in the group will share a birthday by averaging across the 10000 simulation runs. 

```{r}
bday_sim = expand_grid(
    iter = 1:10000,
    n = 2:50) %>% 
  mutate(res = map_lgl(n, bday_sim)) %>% 
  group_by(n) %>% 
  summarize(
    prob = mean(res)
  )

```
Make a plot showing the probability as a function of group size, and comment on your results.
```{r}
ggplot(bday_sim, aes(x = n, y = prob))+
  geom_line()
```

# Problem 2

When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a one-sample t-test.

First set the following design elements:

Fix 𝑛=30
Fix 𝜎=5
Set 𝜇=0
Generate 5000 datasets from the model

𝑥∼𝑁𝑜𝑟𝑚𝑎𝑙[𝜇,𝜎]

For each dataset, save 𝜇̂ and the p-value arising from a test of 𝐻:𝜇=0 using 𝛼=0.05
```{r}
# write a function that generates a normal vector with n = 30, mean = 0, sd = 5
generate_normal_fixed = function(n = 30, mean = 0, sd = 5) {
  norm_vec = rnorm(n = n, mean = mean, sd = sd)
  return (norm_vec)
}
#check that the function works
generate_normal_fixed()

# Generate 5000 datasets, conduct a t-test for each dataset, save the estimate and p-value
fixed_sim_df = tibble(
  iter = 1:5000,
  data = map(1:5000, ~generate_normal_fixed(n = 30, mean = 0, sd = 5))) %>% 
  mutate(
    t_test_res = map(data, \(x) t.test(x, mu = 0)),
    tidy_res = map(t_test_res, tidy)
  ) %>% 
  unnest(tidy_res) %>%
  select(iter, estimate, p.value)

fixed_sim_df
```

Repeat the above for 𝜇={1,2,3,4,5,6}
```{r}
generate_normal = function(n = 30, pop_mean, sd = 5) {
  norm_vec = rnorm(n = n, mean = pop_mean, sd = sd)
  return (norm_vec)
}

sim_df = expand_grid(
    iter = 1:5000,
    pop_mean = c(1,2,3,4,5,6)) %>% 
    mutate(
    data = map(pop_mean, \(x) generate_normal(n = 30, x, sd = 5)),
    t_test_res = map(data, \(x) t.test(x, mu = 0)),
    tidy_res = map(t_test_res, tidy)
  ) %>% 
  unnest(tidy_res) %>% 
  select(iter, pop_mean, estimate, p.value)

sim_df
```
Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of 𝜇on the x axis. Describe the association between effect size and power.
```{r}
# Calculate the power (proportion of p-values < 0.05) for each true mean
power_df = sim_df %>%
  group_by(pop_mean) %>%
  summarize(power = mean(p.value < 0.05))

# Plot the power against the true mean (effect size)
power_df %>%
  ggplot(aes(x = pop_mean, y = power)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Power of the One-Sample t-test vs True Mean",
    x = "True Mean (Effect Size)",
    y = "Power (Proportion of Null Rejected)"
  ) +
  theme_minimal()
```

Power increases as effect size increases. The increase in power is sharp at first as true mean increases and then slows down as power gets close to 1.

Make a plot showing the average estimate of 𝜇̂ on the y axis and the true value of 𝜇on the x axis. Make a second plot (or overlay on the first) the average estimate of 𝜇̂ only in samples for which the null was rejected on the y axis and the true value of 𝜇on the x axis. Is the sample average of 𝜇̂ across tests for which the null is rejected approximately equal to the true value of 𝜇? Why or why not?
```{r}

# calculate average estimates for all samples and for samples where null was rejected
summary_df = sim_df %>%
  group_by(pop_mean) %>%
  summarize(
    avg_estimate_all = mean(estimate),
    avg_estimate_rejected = mean(estimate[p.value < 0.05])
  )

#plot average estimate against the true mean for all and rejected samples
ggplot(summary_df, aes(x = pop_mean)) +
  geom_line(aes(y = avg_estimate_all, color = "All Samples"), size = 1) +
  geom_point(aes(y = avg_estimate_all, color = "All Samples")) +
  geom_line(aes(y = avg_estimate_rejected, color = "Rejected Samples"), size = 1, linetype = "dashed") +
  geom_point(aes(y = avg_estimate_rejected, color = "Rejected Samples")) +
  labs(
    title = "Average Estimate of the Mean vs True Mean",
    x = "True Mean (mu)",
    y = "Average Estimate (mu_hat)",
    color = "Legend"
  ) +
  theme_minimal()

```

The average estimate mu_hat for which the null is rejected is not approximately equal to the true mean. Rather, the average estimates for rejected samples are higher because only samples with larger estimates (further from 0) are more likely to reject the null hypothesis.

# Problem 3
```{r}
# load the homicide data
homicide_raw_df = read_csv("data/homicide-data.csv", na = c("Unknown"))
```

```{r}
# check for missing data
sum(is.na(homicide_raw_df))
summary(homicide_raw_df)
```
The original data contains 52179 rows and 12 columns. Important variables include reported data, victim's first and last name, victim_race <chr>, victim_age <chr> victim_sex <chr>, city, state, latitude and longitude, and disposition <chr>. There is a significant amount of missing data in variables victim_race, victim_age, and victim_sex.

Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).
```{r}
#create a city_state variable
homicide_df = homicide_raw_df %>%
  mutate(city_state = str_c(city, state, sep = ", "))

# Summarizing homicides within each city
homicide_df %>%
  group_by(city_state) %>%
  summarize(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest"))
  )
```
For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.
```{r}
# Subset data for Baltimore, MD
baltimore_data = homicide_df %>%
  filter(city_state == "Baltimore, MD")

# Calculate the number of unsolved homicides and total homicides
unsolved_baltimore = sum(baltimore_data$disposition %in% c("Closed without arrest", "Open/No arrest"))
total_baltimore = nrow(baltimore_data)

# conduct proportion test and save the result as an r object
prop_test_baltimore = prop.test(unsolved_baltimore, total_baltimore)

# Tidy the result & extract estimated proportion and CI
tidy(prop_test_baltimore) %>% 
  select(estimate, conf.low, conf.high)
```
Run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.
```{r}
# For each city, calculate the number of unsolved homicides and total homicides
results_df = 
homicide_df %>%
  group_by(city_state) %>%
  summarize(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest"))
  ) %>%
  mutate(
    prop_test_res = map2(unsolved_homicides, total_homicides, ~prop.test(.x, .y)),
    tidy_res = map(prop_test_res, tidy)
  ) %>%
  unnest(tidy_res) %>%
  select(city_state, estimate, conf.low, conf.high)

results_df
```
Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.
```{r fig.height=8, fig.width=8}
ggplot(results_df, aes(x = reorder(city_state, estimate), y = estimate)) +
  geom_point(color = "blue", size = 2) +  
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +  
  labs(
    x = "City",
    y = "Estimated Proportion of Unsolved Homicides",
    title = "Estimated Proportion of Unsolved Homicides with 95% CI"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
  coord_flip()

ggsave("homicide_plot.png", height = 8, width = 8, units = "in")
```
```{r}
#checking why Tulsa, AL has estimated proportion of 0
homicide_df %>% 
  filter(city_state == "Tulsa, AL") %>% 
  filter(!disposition %in% c("Closed without arrest", "Open/No arrest"))
```
There is only one entry for Tulsa, AL, causing the estimate to be 0.

